Stack Safety is a Security Property

### 2021-04-15 tex/main.tex

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
%  to leave it out of the title but hit it strongly in the abstract.}%
What does ``stack safety'' mean, exactly? The phrase is associated with a
variety of compiler,
run-time, and hardware mechanisms for guarding against corruption of stack
memory.  But these mechanisms typically lack precise specifications,
relying instead on informal descriptions and specific examples of bad
behaviors that they prevent.

We propose formal criteria for stack safety, using the concepts of
language-based security: a combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee'')
and a confidentiality property (``the callee's behavior is insensitive to the
caller's private state'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We argue that these criteria match common intuitions by (1) showing that
they are violated by a variety of intuitively unsafe examples and (2)
using property-based random testing to check that appropriate variants of
the criteria are actually enforced by the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
broken variants).  Specifically, Roessler
and Dehon's ``eager'' micro-policy, which catches violations as early as
possible, enforces a basic ``stepwise'' versions of our confidentiality
and integrity properties. Their more performant ``lazy'' micro-policy
does not, but it does (almost) correspond to weaker, more extensional
variants of the properties.
%
\BCP{I still don't think the next sentence really nails it.  And (sigh) I am
  still not really clear yet on whether our definition applies / might apply
  to other enforcement mechanisms or not.  Saying that our attacker model is
  stronger does not help---that should make it easier for us to describe
  their protections, not harder.  So I don't understand what we're saying
  here.}%
%
While we do not rule out other enforcement mechanisms, our
definitions are strongly informed by micropolicy systems, especially
our attacker model, which is stronger than other enforcement mechanisms
can enable.

\bcp{We should decide between ``criteria'' and ``properties''...}

@@

In the present work, we also distinguish between protecting stack data (like
LSE) and
control flow (like WBCF).  But our formulations of both properties are
somewhat different from those of Skorstengaard et al.  Rather than
defining stack safety implicitly,
as ``what a certain idealized machine does,'' we formalize explicit criteria
based on the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity} and
\emph{stepwise stack confidentiality}.  WBCF is a desireable property, but
we can express safety in an information-flow sense without reference to it,
even safety from attacks that hijack control flow.  \BCP{On consideration
  (especially in light of the weird ``C confuses A by returning directly to
  A instead of returning to B so that B can finish up some stuff before
  returning to A'' example\footnote{Actually, this example is not so weird.
  If C, the attacker, can return a {\em result} directly to A, who is
  expecting a result from B, then that is clearly bad!}), I
agree now that we should discuss WBCF.  It should be kept as distinct as
possible from the ``data safety'' part, since it is not really an
information flow property.  But our treatment of it is interestingly
different from StkTokens, because of our call annotations, and we should
discuss that.}

@@

Our stack-safety criteria depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program {\em annotated} with the locations of
instructions representing calls, and we determine when a return has occurred by tracking
control flow and observing when we reach a {\em return target}. For a given state annotated
as a call, in a standard calling convention, another state fulfils its return target if
it has the same stack pointer and the program counter is at the instruction following
the call, but we abstract over different conventions.
%, returns, and function entry points\BCP{I
%  still want to talk about whether returns and entry points need to be
%  annotated.  To me, these annotations do not make intuitive sense: it is
%  the caller's responsibility to know whether the address that it is calling
%is really an entry point, and we can trust it to check this.  Similarly, it
%is not the callee's role to say when it is returning; it is the caller's
%prerogative to decide when it has been returned to!  I.e., from the caller's
%point of view, it has been returned to when control resumes at the
%instruction following the call with the SP reset to where it was at the
%point of the call. (This is the point where it wants to be able to access
%its own local state again and expected to find it unmolested.)}.
Although call annotations would typically come from a compiler together with
the machine code, we do not assume anything about their provenance.

@@

\bcp{Confusing:}Alternatively, an annotated instruction might not obey the calling convention.
In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
and returns are implemented by fixed sequences of instructions which the tag policy
will force to be executed from start to finish, termed {\em blessed sequences}.
When applying our properties to a micro-policy setting, we will statically rule out some
scenarios with a well-formedness criterion on initial states: all annotated instructions will
be situated at the appropriate instruction of a blessed sequence, tagged according to the policy.
This is all we need for the policy to protect itself: it can then failstop if it would
execute the blessed sequence out of order.

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
%\apt{maybe: ``Stack safety is a testable security property''?}\bcp{That
%  would sacrifice a lot of oomph for a little bit of precision, IMO.  Prefer
%  to leave it out of the title but hit it strongly in the abstract.}%
What does ``stack safety'' mean? The phrase is associated with a
variety of compiler,
run-time, and hardware mechanisms for guarding against corruption of stack
memory.  But these mechanisms typically lack precise specifications,
relying instead on informal descriptions and examples of bad
behaviors that they prevent.

We propose a formal characterization
of stack safety, phrased using concepts from language-based security: a
combination of an integrity property (``the private
state in each caller's stack frame is held invariant by the callee'')
and a confidentiality property (``the callee's behavior is insensitive to the
caller's private state'')\bcp{and something about WBCF...}.
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We use our characterization to validate the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18}.
Specifically, we use property-based random testing to check that Roessler
and Dehon's ``eager'' micro-policy, which catches violations as early as
possible, enforces a simple ``stepwise'' variant of our properties (and
detects several broken variants), and that a slightly repaired version of
their more performant ``lazy'' micro-policy corresponds to weaker and more
extensional variants of the properties.
%
% \BCP{I still don't think the next sentence really nails it.  And (sigh) I am
%   still not really clear yet on whether our definition applies / might apply
%   to other enforcement mechanisms or not.  Saying that our attacker model is
%   stronger does not help---that should make it easier for us to describe
%   their protections, not harder.  So I don't understand what we're saying
%   here.}%
% %
% While we do not rule out other enforcement mechanisms, our
% definitions are strongly informed by micropolicy systems, especially
% our attacker model, which is stronger than other enforcement mechanisms
% can enable.

@@

In the present work, we also distinguish between protecting stack data (like
LSE) and
control flow (like WBCF).  But our formulations of both properties are
somewhat different from those of Skorstengaard et al.  Rather than
defining stack safety implicitly,
as ``what a certain idealized machine does,'' we formalize explicit properties
based on the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity} and
\emph{stepwise stack confidentiality}.  WBCF is a desireable property, but
we can express safety in an information-flow sense without reference to it,
even safety from attacks that hijack control flow.  \BCP{On consideration
  (especially in light of the weird ``C confuses A by returning directly to
  A instead of returning to B so that B can finish up some stuff before
  returning to A'' example\footnote{Actually, this example is not so weird.
  If C, the attacker, can return a {\em result} directly to A, who is
  expecting a result from B, then that is clearly bad!}), I
agree now that we should discuss WBCF.  It should be kept as distinct as
possible from the ``data safety'' part, since it is not really an
information flow property.  But our treatment of it is interestingly
different from StkTokens, because of our call annotations, and we should
discuss that.}

@@

Our stack-safety properties depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program {\em annotated} with the locations of
instructions representing calls, and we determine when a return has occurred by tracking
control flow and observing when we reach a {\em return target}. For a given state annotated
as a call, in a standard calling convention, another state fulfils its return target if
it has the same stack pointer and the program counter is at the instruction following
the call, but we abstract over different conventions.
%, returns, and function entry points\BCP{I
%  still want to talk about whether returns and entry points need to be
%  annotated.  To me, these annotations do not make intuitive sense: it is
%  the caller's responsibility to know whether the address that it is calling
%is really an entry point, and we can trust it to check this.  Similarly, it
%is not the callee's role to say when it is returning; it is the caller's
%prerogative to decide when it has been returned to!  I.e., from the caller's
%point of view, it has been returned to when control resumes at the
%instruction following the call with the SP reset to where it was at the
%point of the call. (This is the point where it wants to be able to access
%its own local state again and expected to find it unmolested.)}.
Although call annotations would typically come from a compiler together with
the machine code, we do not assume anything about their provenance.

@@

\bcp{Confusing, and maybe out of date?}Alternatively, an annotated instruction might not obey the calling convention.
In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
and returns are implemented by fixed sequences of instructions which the tag policy
will force to be executed from start to finish, termed {\em blessed sequences}.
When applying our properties to a micro-policy setting, we will statically rule out some
scenarios with a well-formedness criterion on initial states: all annotated instructions will
be situated at the appropriate instruction of a blessed sequence, tagged according to the policy.
This is all we need for the policy to protect itself: it can then failstop if it would
execute the blessed sequence out of order.
